%! TEX program = xelatex
%% Mauricio Caceres Bravo <mauricio.caceres.bravo@gmail.com>
%% 2023-04-10 11:08

%----------------------------------------------------------------------
\documentclass[11pt]{article}

%\usepackage{nberpreamble}
%\usepackage{changepage}
% \renewcommand\textsc[1]{{\fontfamily{put}\fontshape{sc}\selectfont#1}}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage[skip=\medskipamount]{parskip}
\usepackage[margin=1in]{geometry}

\renewcommand{\arraystretch}{1.25}

\title{\bfseries\scshape Math Camp Syllabus}
\author{\scshape Brown University \\ \scshape Department of Economics}
\date{\scshape Summer 2024}

%\setlength{\parindent}{1.5em}

%----------------------------------------------------------------------
\begin{document}
\maketitle
{
  \bfseries
    \noindent Instructor: Cole Davis \\
    E-mail: cole\_davis@brown.edu \\
    Office Hours: After class and by appointment \\[12pt]
    Times: 8/13, 8/14, 8/15, 8/16, 8/19, 8/20 from 9AM--12:30PM (time includes a 10 min break) \\
    Location: Mencoff Hall (PSTC) 205 \\
    Website: \href{https://cj-davis99.github.io/math-camp.html}{cj-davis99.github.io}
  \vspace{12pt}
}

% ---------------------------------------------------------------------
\section{Course Description}
\label{sec:course_description}

The goal of this course is to present many of the mathematical tools that you will typically encounter throughout your first year at the PhD program.

Keep in mind that this course should serve as a warm-up for the challenging first year that you’ll have. We expect that, at the end of the Math Camp, students will be familiarized with the theory presented and will be able to apply it throughout the first year.

Finally, many of the topics we will cover will be presented in a ``Cookbook'' way, perhaps without the details a rigorous and formal student ideally would want. A more rigorous presentation of the topics we will expose will come shortly (ECON 2010).

% ---------------------------------------------------------------------
\section{Recommended Reading}
\label{sec:recommended_reading}

The course does not follow a particular textbook; hence you are not required to have any. In case you would like to read more about some topics, we suggest the following references: \textit{Real Analysis with Economic Applications} by Efe Ok (henceforth Ok) and \textit{Math for Economists} by Simon and Blume (henceforth SB). In past years these two books (I would argue primarily Ok) have been used in ECON 2010. \textit{Microeconomic Theory} by Mas-Collel, Whinston and Green (henceforth MWG) is the quintessential graduate economics textbook that will be referenced heavily throughout the first year micro theory sequence and contains some useful mathematical appendices. 

Additional reading for introductory mathematical logic: there are plenty of books available that cover this topic, but my personal favorite is Richard Hammack's \textit{Book of Proof} which is freely available \href{https://richardhammack.github.io/BookOfProof/}{here}.

Additional reading for linear algebra: \textit{Linear Algebra Done Right} by Sheldon Axler is arguably the most famous linear algebra textbook and \textit{Linear Algebra} by Jim Hefferon is a solid theory-heavy covering of the material that is freely available \href{https://hefferon.net/linearalgebra/}{here}. 

Additional readings for real analysis: \textit{Principles of Mathematical Analysis} (aka baby Rudin) by Walter Rudin is a classic analysis textbook, but \textit{Analysis I} and \textit{Analysis II} by Terrence Tao are my preferred references and they are freely available for download through Brown's library. 

% ---------------------------------------------------------------------
\section{Homework Assignments}
\label{sec:homework_assignments}

There will be a problem set every two lectures. The assignments will not be collected (though students are free to submit solutions and ask for feedback). The problem sets will consist of basic (and not-so-basic) questions regarding every topic we see. Their goal is to ensure you are aware of (and hopefully comfortable with) the math level that will be required throughout your first year, and to push you towards working with your peers for the first time.

% ---------------------------------------------------------------------
\section{Accommodations for Students with Disabilities}
\label{sec:accommodations_for_students_with_disabilities}

Brown University is committed to full inclusion of all students. Any student with a documented disability is welcome to contact me as early in the summer as possible so that we may arrange reasonable accommodations. As part of this process, please be in touch with Student and Employee Accessibility Services by calling 401-863-9588.

% ---------------------------------------------------------------------
\section{Tentative Schedule}
\label{sec:tentative_schedule}

\begin{table}[h]
    \centering
    \begin{tabular}{c c c c}
         \multicolumn{2}{c}{Lecture} & Topic & References \\ \hline\hline
         \multirow{2}{*}{8/13 (Tue)}& 1st half & Introduction, Mathematical Proofs & Lecture Notes \\
         & 2nd half & Intro to Topology, Limits & Ok A.1–A.3, B.1, C.1 \vspace{.35cm}\\
         \multirow{2}{*}{8/14 (Wed)}& 1st half & Sequences & Ok A.1–A.3, C.1, C.5 \\
         & 2nd half & Continuity & Ok D.1, D.3 \vspace{.35cm}\\
         \multirow{2}{*}{8/15 (Thu)}& 1st half & Compactness, Extreme Value Thm & Ok C.3–C.4, MWG M.I \\
         & 2nd half & Correspondences, Maximum Thm, Fixed Points & Ok E.1–E.3, MWG M.H \vspace{.35cm}\\
         \multirow{2}{*}{8/16 (Fri)}& 1st half & Differentiation, Implicit Function Thm & Ok K.1–K.2, MWG M.E \\
         & 2nd half & Unconstrained Optimization & SB 17, MWG M.D, M.J \vspace{.35cm}\\
         \multirow{2}{*}{8/19 (Mon)}& 1st half & Constrained Optimization, Envelope Thm & SB 18, MWG M.K \\
         & 2nd half & Integration & Ok A.4 \vspace{.35cm}\\
         \multirow{2}{*}{8/20 (Tue)}& 1st half & Linear Algebra & SB 10–11 \\
         & 2nd half & Intro to ODE & SB 24–25
    \end{tabular}
\end{table}

% ---------------------------------------------------------------------
\end{document}

% All-Student Orientation: August 29-30
% International Orientation: August 23-25
% Student of Color Orientation: August 27-28
